---
################################################################################
##                 YOU CANNOT KNIT THIS FILE DIRECTLY FROM HERE!              ##
##              Instead use the england-update-template.R script             ##
################################################################################
output:
  bookdown::pdf_document2: 
    toc: FALSE
fontsize: 12pt
header-includes:
  - \usepackage{xcolor}
  - \usepackage{float}
  - \usepackage{tabu}
  - \usepackage{multirow}
  - \usepackage{colortbl}
  - \usepackage{lscape}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
  - \usepackage{flafter}
params:
  current.year: 2017
  dp.text: 1
  dp.tables: 2
bibliography: england.bib
csl: apa-with-notes.csl
---
  
```{r setup, echo = FALSE, warning=FALSE, message=FALSE}
source(here::here("code/do-not-touch-scripts/england.R"))
```

---
title: "`r paste("Local Authority Parking Finances in England", params$current.year, "-", params$current.year-1999)`"
nocite: |
  `r paste0(bib$refs, collapse = ", ")`
---
  
\renewcommand{\arraystretch}{1.2}

# Introduction


This report uses the MHCLG (formerly DCLG) tables on parking income and expenditure for the period from `r FunFisc(4)` to `r FunFisc()` (`r  eng.bib.main.i.e`). All councils are required to report their financial information to government, in a consistent format, although there may be some differences in interpretation of what that format is and what it covers. Earlier, councils also submitted budgets for the `r FunFisc(-1)` financial year but on a less detailed basis (`r eng.bib.next.budget`). These figures are included where appropriate. The figures have to be on an as-received cash basis without any provision for bad debts or any allowance for capital charges (interest or deprecation).^[The report also includes calculations using data on income, expenditure and budgets for preivious financial years (`r eng.bib.previous.i.e`, `r eng.bib.previous.budget`).]

This report covers `r no.councils` local authorities in England but excludes the national parks which declare parking income. 

The income and cost headings used in this analysis are: 

1. On-street income, expenditure
2. Off-street income, expenditure 
3. Total transport income, expenditure 

The accuracy of this analysis depends on the accuracy with which local authorities complete the annual financial returns they make to central government and upon which this assessment is based. These headings do not always agree with accounts published by local authorities because of their different treatment of overheads and allocation of surpluses to other transport projects.
\newpage

# Summary

```{r sumtab, echo = FALSE, message = FALSE, warning = FALSE}
# print summary data
eng.summary.formatted %>% 
  kable( "latex", 
         booktabs = T, 
         digits = 2, 
         escape = F,
         longtable = T,
         align = c("r"),
         caption = "Summary of parking accounts for England (£ millions)",
         col.names = c("", "", yearz, 
                       paste0(FunFisc(), " budget"),
                       paste0(FunFisc(-1), " budget"),
                       paste0("Change ", current.year, "-", 
                                             current.year-1999, " on ", current.year-1, "-", current.year-1999-1)),
       #  format.args = list(big.mark = ","),
         linesep = "") %>% 
  collapse_rows(columns = 1, latex_hline = "major", valign = "middle") %>% 
  kable_styling(full_width = TRUE, 
                latex_options =c("HOLD_position"), 
                font_size = 9) %>% 
  column_spec(1, width = "1.6cm") %>% 
  column_spec(2, width = "2.3cm") %>% 
  column_spec(3:7, width = "1cm") %>% 
  column_spec(8:9, width = "1.1cm") %>% 
  column_spec(10, width = "1.2cm")  %>% 
  row_spec(c(2, 5), extra_latex_after = "\\cline{2-7} \\cline{10-10}") %>% 
  row_spec(c(8),  extra_latex_after = "\\cline{2-10}") %>% 
  row_spec(9, bold = TRUE) %>% 
  footnote(symbol = "Cells may not add up exactly due to rounding.")
```

Total income from parking in `r FunFisc()` in England was £`r FunDec(pull(summary[5,10]/1000), 2)` billion, `r FunDec(abs(inc.ch), dp.text)` % `r ifelse(inc.ch < 0, "lower", "higher")` than in `r FunFisc(1)`. On-street income `r ifelse(inc.on.ch < 0, "fell", "rose")` by `r FunDec(abs(inc.on.ch), dp.text)` % and off-street income `r ifelse(sign(inc.on.ch) == sign(inc.off.ch), "", ifelse(inc.off.ch < 0, "fell", "rose"))` by `r FunDec(abs(inc.off.ch), dp.text)` % . Expenditure `r ifelse(exp.on.ch < 0, "fell", "rose")` by `r FunDec(abs(exp.on.ch), dp.text)` % for on-street and `r ifelse(sign(exp.on.ch) == sign(exp.off.ch), "", ifelse(exp.off.ch < 0, "fell", "rose"))` by `r FunDec(abs(exp.off.ch), dp.text)` % for off-street parking. As a result, the combined surplus from both on- and off-street parking was £`r FunDec(pull(summary[5,12]), 0)` million, `r FunDec(abs(sur.ch), dp.text)`% `r ifelse(sur.ch < 0, "lower", "higher")`  than the previous financial year and `r FunDec(abs(sur.ch4), dp.text)`% `r ifelse(sur.ch4 < 0, "lower", "higher")`  than four years ago. 

Local authorities `r ifelse(sur.budg.diff > 0, "exceeded", "fell short of")` their `r FunFisc()` budgeted surplus of £`r FunDec(summary$surplus.total[6], 0)` million by `r FunDec(abs(sur.budg.diff), dp.text)` %. Their budgeted surplus for this next financial year (`r FunFisc(-1)`) is £`r FunDec(summary$surplus.total[7], 0)` million, `r FunDec(abs(sur.budg.inc), dp.text)`% `r ifelse(sur.budg.inc < 0, "lower", "higher")`   than the actual for `r FunFisc()`. If the `r FunFisc(-1)` budget `r ifelse(sur.budg.diff > 0, "is exceeded", "falls short")` by the same proportion as the `r FunFisc()` outcome, then total parking surpluses would amount to £`r FunDec(sur.hyp, 0)` million.

The on-street parking surpluses must by law be used for transport projects. Overall, local transport net expenditure in `r FunFisc()` `r ifelse(abs(tran.ch) < 0.5, "remained relatively unchanged at", ifelse(tran.ch < 0, "fell to", "increased to"))`  £`r FunDec(summary$transport.total[5]/1000, 1)` billion. Without the contribution from parking surpluses, net expenditure would have been `r FunDec(summary$prop.net.expen[5], dp.text)`% lower.

\newpage
```{r sumtablondon, echo = FALSE, message = FALSE, warning = FALSE}
# print summary data
eng.summary.london.formatted %>% 
  kable( "latex", 
         booktabs = T, 
         digits = 2, 
         escape = F,
         longtable = T,
         align = c("r"),
         caption = "Summary of parking accounts for England by region (£ millions)",
         col.names = c("", "", "London", "Rest of England", "All of England",
                       "London percentage"),
         linesep = "") %>% 
  collapse_rows(columns = 1, latex_hline = "major", valign = "middle") %>% 
  kable_styling(full_width = TRUE, 
                latex_options =c("HOLD_position"), 
                font_size = 10) %>% 
  column_spec(1, width = "2cm") %>% 
  column_spec(2, width = "2cm")  %>% 
    row_spec(c(2,5, 8),  extra_latex_after = "\\cline{2-6}") 
```


```{r fig1, echo = FALSE, fig.cap = "Parking revenues--England", fig.pos = "H", fig.showtext=TRUE,}
old.par <- par(mar = c(5.1, 4.1, 3.1, 2.1))
plot(eng.plot$year, eng.plot$income.total, 
     ylim = c(0, max(eng.plot$income.total, na.rm = TRUE) ), 
     type = "l",
     lwd = 2,
     col = viridis(3)[1],
     bty = "n",
     xlab = "Fiscal Year",
     ylab = "£ (millions)",
     axes = FALSE, family = "meri", cex.lab = 0.8)
axis(1, at = eng.plot$year, 
     labels = rep("", length(eng.plot$year),
                  family = "meri", cex.axis = 0.6, las = 2, srt = 45))
text(eng.plot$year, par("usr")[3] - 0.5, 
     labels = paste0(eng.plot$year, "-", eng.plot$year-1999), 
     srt = 45, adj = c(1.2,1.5), xpd = TRUE, family = "meri", cex = 0.6)
axis(2, las = 2, family = "meri", cex.axis = 0.6 )
grid(nx = NA, ny = NULL, lty = "93")
lines(eng.plot$year, eng.plot$surplus.budget, 
      type = "l",
      lwd = 2, lty = "63",
      col = viridis(3)[3])
lines(eng.plot$year, eng.plot$expend.total, 
      type = "l",
      lwd = 2,
      col = viridis(3)[2])
lines(eng.plot$year, eng.plot$surplus.total, 
      type = "l",
      lwd = 2,
      col = viridis(3)[3])

par(family = "meri")
legend("bottomright",
       legend = c("Income", "Expenditure", "Surplus", "Budgeted surplus"),
       bty = "n",
       col = viridis(3)[c(1,2,3,3)], 
       lwd = 2, cex = 0.8, lty = c("99","99","99","63"), xpd = TRUE)
par(old.par)
```

\renewcommand{\arraystretch}{1.2}


```{r compare, echo = FALSE}
# pring comparison table for all three countrires
sum.gb.formatted %>% 
  kable( "latex", 
         booktabs = T, 
         digits = c(rep(0, 6), 2), 
         escape = F,
         longtable = T,
         align = c("r"),
         col.names = c("", "England without London", "London", 
                       "Scotland", "Wales", "Great Britain"),
         caption = paste0("Comparison of parking income and expenditure across the nations of Great Britain (£ millions, latest year available)"), 
         format.args = list(big.mark = ","),
         linesep = "") %>% 
  kable_styling(full_width = TRUE, 
                latex_options =c("hold_position", "striped", "repeat_header"), 
                font_size = 10) %>% 
  column_spec(1, width = "5cm") %>% 
  row_spec(c(1, 4), hline_after = TRUE)
```

Table \@ref(tab:compare) provides a comparison with London, England excluding London, Wales, and Scotland for the most recent available data, while Table \@ref(tab:change) compares the changes between `r FunFisc()` and the previous year, with the average annual change over the four-year period starting in `r FunFisc(4)` (or the most recent four-year period for which data is available). In the last year the surpluses for London have `r ifelse(lond.ann < 0, "decreased", "increased")` by `r FunDec(abs(lond.ann), dp.text)` %, which is `r ifelse(lond.ann < lond.4av, "less", "more")` than the average annual increase observed over the preceding four years, which was `r FunDec(lond.4av, dp.text)` %. For the rest of England excluding London the surpluses have  `r ifelse(rest.ann < 0, "decreased", "increased")` by `r FunDec(abs(rest.ann), dp.text)` %, which is `r ifelse(rest.ann < rest.4av, "less", "more")` than the four-year average annual increase at `r FunDec(rest.4av, dp.text)` %.

On average, parking surpluses in Great Britain have risen by about `r FunDec(gb.ann, dp.text)` % annually over the four years compared with `r FunDec(rpi.annual.gb, dp.text)` % annually for the Retail Prices Index during the same period (`r eng.bib.rpi`).`r ifelse(gb.mr.year != current.year,  paste0("^[The most recent data available for Great Britain as a whole is from the year ", FunFisc(current.year-gb.mr.year), " and all calculations are therefore performed for the four years previous i.e. from ", FunFisc(current.year-gb.mr.year+4), ".]"), "")`


```{r change}
# print change over 5 years. 
sum.gb.change.tab.formatted  %>% 
  kable( "latex", 
         booktabs = T, 
         digits = c(rep(0, 6), 2), 
         escape = F,
         longtable = T,
         align = c("r"),
         col.names = c("",  "England without London", "London", 
                       "Scotland", "Wales", "Great Britain"),
         caption = paste0("Changes in parking income and expenditure over previous four years (from most recent year available) across the nations of Great Britain"), 
         format.args = list(big.mark = ","),
         linesep = "") %>% 
  kable_styling(full_width = TRUE, 
                latex_options =c("HOLD_position", "striped", "repeat_header"), 
                font_size = 9) %>% 
  column_spec(1, width = "6cm") %>% 
  # column_spec(6, width = "1.8cm") %>% 
  row_spec(c(1,3,5), hline_after = TRUE)
```

\newpage

# Income 

Parking income is derived from three main sources: meter income, residents’ and  business permits, and penalties. Other sources are towing and storage charges although relatively few councils operate pounds now. Total income in `r FunFisc()` was £`r formatC(inc.tot, big.mark = ",")` million, `r FunDec(inc.ch, dp.text)`% higher than in `r FunFisc(1)`. In London boroughs the total income was £`r ifelse(inc.london.tot >= 1000, paste(FunDec(inc.london.tot/1000, 0), "billion"), paste(FunDec(inc.london.tot, 0), "million"))`, `r FunDec(inc.london.tot.ch, dp.text)` % more than in the previous year, while the total for the rest of England was £`r ifelse(inc.rest.tot >= 1000, paste(FunDec(inc.rest.tot/1000, 1), "billion"), paste(FunDec(inc.rest.tot, 0), "million"))`, `r FunDec(inc.rest.tot.ch, dp.text)` % more than in the previous year. 

Table \@ref(tab:incometablondon) shows the incomes for all the London boroughs for the previous 5 years, their most recent annual change and the average annual change in income over the preceding 5 years. Table \@ref(tab:incometabrest) has the same information for the top 20 councils outside of London. For a full table of all the local authorities in London see Table \@ref(tab:incometab)  in the Appendix. 


```{r incometablondon, echo = FALSE, message = FALSE, warning = FALSE}
eng.income.london.formatted %>% 
  kable( "latex", 
         booktabs = T, 
         digits = c(rep(dp.tables, 6), 2), 
         escape = F,
         longtable = T,
         align = c("r"),
         caption = "Parking income for London boroughs (£ millions)",
         format.args = list(big.mark = ","),
         linesep = "",
         col.names = c("\\multirow{1}{*}[0pt]{Local Authority}", yearz, 
                       paste0("Change ", FunFisc(), 
                              " on ", FunFisc(1)),
                       paste0("Average annual change since ", FunFisc(4)))) %>% 
  kable_styling(full_width = TRUE, 
                latex_options =c("hold_position", "striped", "repeat_header"), 
                font_size = 9) %>% 
  column_spec(1, width = "4.5cm") %>% 
  column_spec(7, width = "1.5cm") %>% 
  column_spec(8, width = "1.5cm") %>% 
  row_spec(c(33, 35), hline_after = TRUE)

```


In England as a whole `r FunFirstup(FunN2W(income.bin["poz"]))` `r ifelse(income.bin["poz"] == 1, "council", "councils")` increased their income over the past year and `r FunN2W(income.bin["neg"])` decreased their income. `r if(!is.na(income.bin["zero"])) paste("Another", ifelse ( income.bin["zero"] == 1, "single",FunN2W(income.bin["zero"])), ifelse ( income.bin["zero"] == 1, "council's income", "councils' incomes"), "stayed the same.")`
 `r if(!is.na(income.bin["na"])) paste(ifelse(income.bin["na"] == 1, paste("There is", FunN2W(income.bin["na"]), "council"), paste("There are", FunN2W(income.bin["na"]), "councils")), "that did not report income data in this or the previous fiscal year, so we cannot define the change.")`

The top three English councils by income were `r paste0(names(eng.income.top3)[1], ", ", names(eng.income.top3)[2], ", and ", names(eng.income.top3)[3])`, and between them accounted for `r FunDec(eng.income.top3[1], dp.text)` % of parking income. 

The three councils which had the largest year-on-year percentage increases in income were `r eng.income.change.top3[1,2]` (up `r FunDec(eng.income.change.top3$change[1], dp.text)` %), `r eng.income.change.top3[2,2]` (up `r FunDec(eng.income.change.top3$change[2], dp.text)` %) and `r eng.income.change.top3[3,2]` (up `r FunDec(eng.income.change.top3$change[3], dp.text)` %).
`r if(nrow(eng.income.excluded.top) > 0) paste0(FunMultiText(eng.income.excluded.top$auth.name), " showed a larger percentage change but parking income was under £30,000.")` The biggest decreases were in `r eng.income.change.bottom2[2,2]` (`r FunDec(abs(eng.income.change.bottom2$change[2]), dp.text)` %), and `r eng.income.change.bottom2[1,2]` (`r FunDec(abs(eng.income.change.bottom2$change[1]), dp.text)`  %)`r if(nrow(eng.income.excluded.bottom) > 0) paste0(" (excluding ", FunMultiText(eng.income.excluded.bottom$auth.name), " where income was under £30,000)")`.



```{r incometabrest, echo = FALSE, message = FALSE, warning = FALSE}
eng.income.rest.formatted %>% 
  slice(1:20, (n()-2):n()) %>% 
  kable( "latex", 
         booktabs = T, 
         digits = c(rep(dp.tables, 6), 2), 
         escape = F,
         longtable = T,
         align = c("r"),
         caption = "Parking income for the rest of England outside London, top 20 councils (£ millions)",
         format.args = list(big.mark = ","),
         linesep = "",
         col.names = c("\\multirow{1}{*}[0pt]{Local Authority}", yearz, 
                       paste0("Change ", FunFisc(), 
                              " on ", FunFisc(1)),
                       paste0("Average annual change since ", FunFisc(4)))) %>% 
  kable_styling(full_width = TRUE, 
                latex_options =c("hold_position", "repeat_header"), 
                font_size = 9) %>% 
  column_spec(1, width = "4.5cm") %>% 
  column_spec(7, width = "1.5cm") %>% 
  column_spec(8, width = "1.5cm") %>% 
  row_spec(c(20, 22), hline_after = TRUE)

```



# Expenditure

Overall expenditure on parking `r ifelse( exp.ch > 0, "rose", "fell")` to £`r FunDec(exp.tot,0)` million, `r ifelse( exp.ch > 0, "up", "down")` `r FunDec(abs(exp.ch) , dp.text)` % from `r FunFisc(1)` for both on-street and off-street. Expenditure in London went `r ifelse( exp.london.tot.ch > 0, "up", "down")` by `r FunDec(abs(exp.london.tot.ch), dp.text)` % and `r ifelse(sign(exp.london.tot.ch == exp.rest.tot.ch), "", ifelse(exp.rest.tot.ch > 0, "up", "down"))` by `r FunDec(abs(exp.rest.tot.ch), dp.text)` % in the rest of England. Expenditure represents `r FunDec(exp.prop.inc.on, dp.text)` % of income for on-street parking and `r FunDec(exp.prop.inc.off, dp.text)` % for off-street. 

Table \@ref(tab:expendtablondon) ranks the London boroughs in terms of expenditure on parking while Table  \@ref(tab:expendtabrest) lists the top 20 English councils outside London. See Table \@ref(tab:expendtab) in the Appendix for a comprehensive list of all 353 local authorities in England. 


Overall expenditure has `r ifelse( eng.expend.t[2] > 0, "risen", "fallen")` by £`r ifelse(abs(eng.expend.t[2]) <= 0.05, paste(FunDec(abs(eng.expend.t[2]*1000), 0), "thousand"), paste(FunDec(abs(eng.expend.t[2]), 0), "million"))` (`r FunDec(abs(eng.expend.t[1]), dp.text)` %) after a `r ifelse( eng.expend.t[3] > 0, ifelse(abs(eng.expend.t[3]) <= 0.05, paste0("rise of £", FunDec(eng.expend.t[3]*1000, 0), " thousand last year"), paste0("rise of £", FunDec(eng.expend.t[3], 0), " million last year")), ifelse(abs(eng.expend.t[3]) <= 0.05, paste0("fall of £", FunDec(eng.expend.t[3]*1000, 0), " thousand last year"), paste0("fall of £", FunDec(eng.expend.t[3], 0), " million last year")))`.


In England as a whole `r FunFirstup(FunN2W(expend.bin["poz"]))` `r ifelse(expend.bin["poz"] == 1, "council", "councils")` increased their expenditures over the past year and `r FunN2W(expend.bin["neg"])` decreased them. `r if(!is.na(expend.bin["zero"])) paste("Another", ifelse ( expend.bin["zero"] == 1, "single",FunN2W(expend.bin["zero"])), ifelse ( expend.bin["zero"] == 1, "council's expenditure", "councils' expenditures"), "stayed the same.")` `r if(!is.na(expend.bin["na"])) paste(ifelse(expend.bin["na"] == 1, paste("There is", FunN2W(expend.bin["na"]), "council"), paste("There are", FunN2W(expend.bin["na"]), "councils")), "that did not report expenditure data in this or the previous fiscal year, so we cannot define the change.")`


The top three English councils by expenditure `r paste0(names(eng.expend.top3)[1], ", ", names(eng.expend.top3)[2], ", and ", names(eng.expend.top3)[3])`, and between them accounted for `r FunDec(eng.expend.top3[1], dp.text)` % of parking expenditure in the country.  

The largest increase in expenditure in England occured in `r eng.expend.change.top3[1,2]` where it increased by `r FunDec(eng.expend.change.top3$change[1], dp.text)` %
while `r eng.expend.change.top3[2,2]` and `r eng.expend.change.top3[3,2]` increased  by `r FunDec(eng.expend.change.top3$change[2], dp.text)` % and `r FunDec(eng.expend.change.top3$change[3], dp.text)` % respectively`r if(nrow(eng.expend.excluded.top) > 0) paste0(" (excluding ", FunMultiText(eng.expend.excluded.top$auth.name), " with expenditure under £30,000)")`. The biggest decreases were in `r eng.expend.change.bottom2[2,2]` (`r FunDec(abs(eng.expend.change.bottom2$change[2]), dp.text)` %), and `r eng.expend.change.bottom2[1,2]` (`r FunDec(abs(eng.expend.change.bottom2$change[1]), dp.text)`  %)`r if(nrow(eng.expend.excluded.bottom) > 0) paste0(" (excluding ", FunMultiText(eng.expend.excluded.bottom$auth.name), " with expenditure under £30,000)")`.


```{r expendtablondon, echo = FALSE, message = FALSE, warning = FALSE}
eng.expend.london.formatted %>% 
  kable( "latex", 
         booktabs = TRUE, 
         digits = dp.tables, 
         escape = FALSE,
         longtable = TRUE,
         align = c("r"),
         caption = "Parking expenditure for London boroughs (£ millions)",
         format.args = list(big.mark = ","),
         linesep = "",
         col.names = c("\\multirow{1}{*}[0pt]{Local Authority}", yearz, 
                       paste0("Change ", FunFisc(), 
                              " on ", FunFisc(1)), 
                       paste0("Average annual change since ", FunFisc(4)),
                       "Expenditure as proportion of income")) %>% 
  kable_styling(full_width = TRUE, 
                latex_options =c("HOLD_position", "striped", "repeat_header"), 
                font_size = 9) %>% 
  column_spec(1, width = "4.2cm") %>% 
    column_spec(7, width = "1.4cm") %>%
    column_spec(8, width = "1.4cm") %>%
  column_spec(9, width = "1.4cm") %>% 
  row_spec(c(33,35), hline_after = TRUE)
```



```{r expendtabrest, echo = FALSE, message = FALSE, warning = FALSE}
eng.expend.rest.formatted %>% 
  slice(1:20, (n()-2):n()) %>% 
  kable( "latex", 
         booktabs = TRUE, 
         digits = dp.tables, 
         escape = FALSE,
         longtable = TRUE,
         align = c("r"),
         caption = "Parking expenditure for the rest of England outside London, top 20 councils (£ millions)",
         format.args = list(big.mark = ","),
         linesep = "",
         col.names = c("\\multirow{1}{*}[0pt]{Local Authority}", yearz, 
                       paste0("Change ", FunFisc(), 
                              " on ", FunFisc(1)), 
                       paste0("Average annual change since ", FunFisc(4)),
                       "Expenditure as proportion of income")) %>% 
  kable_styling(full_width = TRUE, 
                latex_options =c("HOLD_position", "striped", "repeat_header"), 
                font_size = 9) %>% 
  column_spec(1, width = "4.2cm") %>% 
    column_spec(7, width = "1.4cm") %>%
    column_spec(8, width = "1.4cm") %>%
  column_spec(9, width = "1.4cm") %>% 
  row_spec(c(20,22), hline_after = TRUE)
```


Table \@ref(tab:expendprop) shows the proportion of income taken up by expenditure over the previous four years for London and table \@ref(tab:expendprop2) for the top 20 councils outside London. For data for all English councils, see Table \@ref(tab:expendprop3) in the Appendix. 

```{r expendprop}
eng.expend.props.london.formatted %>% 
  kable( "latex", 
         booktabs = TRUE, 
         escape = FALSE,
         longtable = TRUE,
         align = c("r"),
         caption = "Parking expenditure as proportion of parking income for London boroughs",
         format.args = list(big.mark = ","),
         linesep = "",
         col.names = c("\\multirow{1}{*}[0pt]{Local Authority}", yearz)) %>% 
  kable_styling(full_width = TRUE, 
                latex_options =c("HOLD_position", "striped", "repeat_header"), 
                font_size = 9) %>% 
  column_spec(1, width = "4.2cm") %>% 
    column_spec(7, width = "1.5cm") %>%
    column_spec(8, width = "1.5cm") %>%
  column_spec(9, width = "1.5cm") %>% 
  row_spec(c(33, 35), hline_after = TRUE) %>% 
  footnote(symbol = "Any empty cells indicate the council reported no income and/or no expenditure")
```


```{r expendprop2}
eng.expend.props.rest.formatted %>% 
  slice(1:20, (n()-2):n()) %>% 
  kable( "latex", 
         booktabs = TRUE,
         escape = FALSE,
         longtable = TRUE,
         align = c("r"),
         caption = "Parking expenditure as proportion of parking income for the rest of England, top 20 councils",
         format.args = list(big.mark = ","),
         linesep = "",
         col.names = c("\\multirow{1}{*}[0pt]{Local Authority}", yearz)) %>% 
  kable_styling(full_width = TRUE, 
                latex_options =c("HOLD_position", "striped", "repeat_header"), 
                font_size = 9) %>% 
  column_spec(1, width = "4.2cm") %>% 
  row_spec(c(20, 22), hline_after = TRUE) %>% 
  footnote(symbol = "Any empty cells indicate the council reported no income and/or no expenditure")
```

\newpage
# Surpluses 

In total, councils in England made a surplus of £`r FunDec(sur.tot,0)` million in `r FunFisc()`, `r FunDec(abs(sur.ch), dp.text)` % `r ifelse(sur.ch > 0, "higher", "lower")` than in `r FunFisc(1)`. Of this total, £`r FunDec(sur.london, 0)` million (`r FunDec(sur.london.prop, dp.text)` %) came from London and £`r FunDec(sur.rest, 0)` million from the rest of England. This excludes a £`r FunDec(surplus.nat.parks/1000, dp.text)` million surplus from parking in National Parks. Money from any workplace parking levy schemes is excluded from the analysis for the sake of clear comparison between local authorities.  It is currently unique to Nottingham.


The total *surpluses* for London amounted to £`r FunDec((london.surplus.totals.table.sur), dp.text)` million between `r  london.poz.surplus ` authorities, of which the top three--`r paste0(names(london.surplus.top3)[1], ", ", names(london.surplus.top3)[2], ", and ", names(london.surplus.top3)[3])`--accounted for `r FunDec(london.surplus.top3[1],  dp.text)` %.  

`r ifelse(is.na(surplus.bin["neg"]), "No London councils made a loss last year", paste(ifelse( surplus.bin["neg"] == 1, paste(FunFirstup(FunN2W(surplus.bin["neg"])) , "council"), paste("In London", FunFirstup(FunN2W(surplus.bin["neg"])) , "councils")), "made a loss with the total of parking deficits",  ifelse(london.surplus.totals.table.def > london.surplus.totals.table.def.last, "falling", "rising"), "to £", FunDec(-london.surplus.totals.table.def, dp.text), "million from £",  FunDec(-london.surplus.totals.table.def.last, dp.text), "million last year"))` (see Table \@ref(tab:surlondtab)).

Overall, parking contributed £`r FunDec(london.surplus.totals.table.tot, dp.text)` million to local authority finances in London in `r FunFisc()` compared with £`r FunDec(london.surplus.totals.table.tot.last, dp.text)` million in `r FunFisc()`, an increase of `r FunDec(london.surplus.totals.table.tot.ch, dp.text)` %. 

`r london.surplus.change.top3[1,2]` increased their surplus by `r FunDec(london.surplus.change.top3$change[1], dp.text)` %
while `r london.surplus.change.top3[2,2]` and `r london.surplus.change.top3[3,2]` increased  by `r FunDec(london.surplus.change.top3$change[2], dp.text)` % and `r FunDec(london.surplus.change.top3$change[3], dp.text)` % respectively`r if(nrow(london.surplus.excluded.top) > 0) paste0(" (excluding ", FunMultiText(london.surplus.excluded.top$auth.name), " with a surplus under £30,000)")`. The biggest decreases were in `r london.surplus.change.bottom2[2,2]` (`r FunDec(abs(london.surplus.change.bottom2$change[2]), dp.text)` %), and `r london.surplus.change.bottom2[1,2]` (`r FunDec(abs(london.surplus.change.bottom2$change[1]), dp.text)`  %)`r if(nrow(london.surplus.excluded.bottom) > 0) paste0(" (excluding ", FunMultiText(london.surplus.excluded.bottom$auth.name), " with a surplus under £30,000)")`.

`r ifelse(nrow(london.deficit.valid) == 0, "None of the London boroughs has had a deficit two years in a row.", paste0(london.deficit.change.top3[1,2], " increased their deficits by ", FunDec(london.deficit.change.top3$change[1], dp.text)," % while ", london.deficit.change.top3[2,2], " and ",  london.deficit.change.top3[3,2], " increased  by ", FunDec(london.deficit.change.top3$change[2], dp.text), " % and ", FunDec(london.deficit.change.top3$change[3], dp.text), " % respectively ",  if(nrow(london.deficit.excluded.top) > 0) paste0(" (excluding ", FunMultiText(london.deficit.excluded.top$auth.name), " with a deficit under £30,000)"), " The biggest decreases were in ",  london.deficit.change.bottom2[2,2], " (", FunDec(abs(london.deficit.change.bottom2$change[2]), dp.text), " %), and ", london.deficit.change.bottom2[1,2], "  (",  FunDec(abs(london.deficit.change.bottom2$change[1]), dp.text), " %)", if(nrow(london.deficit.excluded.bottom) > 0) paste0(" (excluding ", FunMultiText(london.deficit.excluded.bottom$auth.name), " with a deficit under £30,000)")))`



```{r surlondtab}
# print summary data
london.surplus.totals.table.formatted %>% 
  kable("latex", 
         booktabs = T, 
         digits = 0, 
         longtable = T,
         escape = F,
         align = c("r"),
         caption = "Parking surpluses (and deficits) for London (£ millions)",
       # format.args = list(big.mark = ","),
         linesep = "",
         col.names = c("\\multirow{1}{*}[0pt]{Local Authority}", yearz, 
                       paste0("Change ", current.year, "-", current.year-1999, 
                              " on ", current.year-1, "-", current.year-1999-1))) %>% 
  kable_styling(full_width = TRUE, 
                latex_options =c("HOLD_position", "striped", "repeat_header"), 
                font_size = 9) %>% 
  column_spec(1, width = "4.5cm") %>% 
  row_spec(c(33, 35), hline_after = TRUE) %>% 
  footnote(symbol = "Where the change in surplus is actually a change in deficit, the values are in italics")
```

```{r surengtab}
# print surpluses for non london councils 
eng.rest.surplus.totals.table.formatted %>% 
  kable("latex", 
         booktabs = T, 
         digits = 0, 
         longtable = T,
         escape = F,
         align = c("r"),
         caption = "Parking surpluses (on-street and off-street) for top 20 councils outside London (£ millions)",
         linesep = "",
         col.names = c("\\multirow{1}{*}[0pt]{Local Authority}", yearz, 
                       paste0("Change ", current.year, "-", current.year-1999, 
                              " on ", current.year-1, "-", current.year-1999-1))) %>% 
  kable_styling(full_width = TRUE, 
                latex_options =c("HOLD_position", "striped", "repeat_header"), 
                font_size = 9) %>% 
  column_spec(1, width = "4.2cm") %>% 
  row_spec(c(20, 22), hline_after = TRUE) 
```


The total surplus from parking outside London of £`r FunDec(sur.rest,0)` million in `r FunFisc()` was  `r FunDec(abs(sur.rest.tot.ch), dp.text)` % `r ifelse(sur.rest.tot.ch > 0, "higher", "lower")` than in  `r FunFisc(1)`. `r sur.rest.top.la` was the highest earning council outside London, ranking as the `r ordinal(sur.rest.top.rank)` largest  nationally, and with a surplus in `r FunFisc()` of £`r FunDec(sur.rest.top.amount, dp.text)` million, `r FunDec(abs(sur.rest.top.ch), dp.text)` % `r ifelse(sur.rest.top.ch > 0, "higher", "lower")` than in `r FunFisc(1)`. Table \@ref(tab:surengtab) lists the top 20 councils outside of London. The full table for all English councils is available in the Appendix as Table \@ref(tab:surrestfull) . 

The total *surpluses* outside London amounted to £`r FunDec((rest.surplus.totals.table.tot), dp.text)` million between `r  rest.poz.surplus ` authorities, of which the top three--`r paste0(names(rest.surplus.top3)[1], ", ", names(rest.surplus.top3)[2], ", and ", names(rest.surplus.top3)[3])`--accounted for `r FunDec(rest.surplus.top3[1],  dp.text)` %.  `r FunFirstup(FunN2W(surplus.bin.rest["neg"]))` `r ifelse(surplus.bin.rest["neg"] == 1, "council", ifelse(surplus.bin.rest["neg"] == 0, "No councils", "councils"))` made a loss with the total of parking deficits `r ifelse(rest.surplus.totals.table.def > rest.surplus.totals.table.def.last, "falling", "rising")` to £`r FunDec(-rest.surplus.totals.table.def, dp.text)` million from £`r FunDec(-rest.surplus.totals.table.def.last, dp.text)` million last year. 


`r rest.surplus.change.top3[1,2]` increased their surplus by `r FunDec(rest.surplus.change.top3$change[1], dp.text)` %
while `r rest.surplus.change.top3[2,2]` and `r rest.surplus.change.top3[3,2]` increased  by `r FunDec(rest.surplus.change.top3$change[2], dp.text)` % and `r FunDec(rest.surplus.change.top3$change[3], dp.text)` % respectively`r if(nrow(rest.surplus.excluded.top) > 0) paste0(" (excluding ", FunMultiText(rest.surplus.excluded.top$auth.name), " with a surplus under £30,000)")`. The biggest decreases were in `r rest.surplus.change.bottom2[2,2]` (`r FunDec(abs(rest.surplus.change.bottom2$change[2]), dp.text)` %), and `r rest.surplus.change.bottom2[1,2]` (`r FunDec(abs(rest.surplus.change.bottom2$change[1]), dp.text)`  %)`r if(nrow(rest.surplus.excluded.bottom) > 0) paste0(" (excluding ", FunMultiText(rest.surplus.excluded.bottom$auth.name), " with a surplus under £30,000)")`.

`r ifelse(nrow(rest.deficit.valid) == 0, "None of the councils has had a deficit two years in a row", paste0(rest.deficit.change.top3[1,2], " increased their deficits by ", FunDec(rest.deficit.change.top3$change[1], dp.text)," % while ", rest.deficit.change.top3[2,2], " and ",  rest.deficit.change.top3[3,2], " increased  by ", FunDec(rest.deficit.change.top3$change[2], dp.text), " % and ", FunDec(rest.deficit.change.top3$change[3], dp.text), " % respectively ",  if(nrow(rest.deficit.excluded.top) > 0) paste0(" (excluding ", FunMultiText(rest.deficit.excluded.top$auth.name), " with a deficit under £30,000)"), " The biggest decreases were in ",  rest.deficit.change.bottom2[2,2], " (", FunDec(abs(rest.deficit.change.bottom2$change[2]), dp.text), " %), and ", rest.deficit.change.bottom2[1,2], "  (",  FunDec(abs(rest.deficit.change.bottom2$change[1]), dp.text), " %)", if(nrow(rest.deficit.excluded.bottom) > 0) paste0(" (excluding ", FunMultiText(rest.deficit.excluded.bottom$auth.name), " with a deficit under £30,000)")))`

\newpage


\newpage
# Appendix 

## Income data

\renewcommand{\arraystretch}{1.2}

```{r incometab, echo = FALSE, message = FALSE, warning = FALSE}
eng.income.formatted %>% 
  kable( "latex", 
         booktabs = T, 
         digits = c(rep(0, 6), 2), 
         escape = F,
         longtable = T,
         align = c("r"),
         caption = "Parking income for England (£,000)",
         format.args = list(big.mark = ","),
         linesep = "",
         col.names = c("\\multirow{1}{*}[0pt]{Local Authority}", yearz, 
                       paste0("Change ", FunFisc(), 
                              " on ", FunFisc(1)),
                       paste0("Average annual change since ", FunFisc(4)))) %>% 
  kable_styling(full_width = TRUE, 
                latex_options =c("hold_position", "striped", "repeat_header"), 
                font_size = 8) %>% 
  column_spec(1, width = "4.2cm") %>% 
  column_spec(7, width = "1.5cm") %>% 
  column_spec(8, width = "1.5cm") %>% 
  row_spec(c(353, 355), hline_after = TRUE)

```

\newpage
## Expenditure data

```{r expendtab, echo = FALSE, message = FALSE, warning = FALSE}
eng.expend.formatted %>% 
  kable( "latex", 
         booktabs = TRUE, 
         digits = 0, 
         escape = FALSE,
         longtable = TRUE,
         align = c("r"),
         caption = "Parking expenditure(£,000)",
         format.args = list(big.mark = ","),
         linesep = "",
         col.names = c("\\multirow{1}{*}[0pt]{Local Authority}", yearz, 
                       paste0("Change ", FunFisc(), 
                              " on ", FunFisc(1)), 
                       paste0("Average annual change since ", FunFisc(4)),
                       "Expenditure as proportion of income")) %>% 
  kable_styling(full_width = TRUE, 
                latex_options =c("HOLD_position", "striped", "repeat_header"), 
                font_size = 8) %>% 
  column_spec(1, width = "3.9cm") %>% 
    column_spec(7, width = "1.5cm") %>%
    column_spec(8, width = "1.5cm") %>%
  column_spec(9, width = "2cm") %>% 
  row_spec(c(353,355), hline_after = TRUE)
```


\newpage

```{r expendprop3}

eng.expend.props.formatted %>% 
  kable( "latex", 
         booktabs = TRUE, 
         digits = 0, 
         escape = FALSE,
         longtable = TRUE,
         align = c("r"),
         caption = "Parking expenditure as proportion of parking income for all of England",
         format.args = list(big.mark = ","),
         linesep = "",
         col.names = c("\\multirow{1}{*}[0pt]{Local Authority}", yearz)) %>% 
  kable_styling(full_width = TRUE, 
                latex_options =c("HOLD_position", "striped", "repeat_header"), 
                font_size = 8) %>% 
  column_spec(1, width = "3.9cm") %>% 
  row_spec(c(353, 355), hline_after = TRUE) %>% 
  footnote(symbol = "Any empty cells indicate the council reported no income and/or no expenditure")

```

\newpage
## Surpluses

```{r surrestfull}
# print surpluses for non london councils 
rest.surplus.totals.table.full.formatted %>% 
  kable("latex", 
         booktabs = T, 
         digits = 0, 
         longtable = T,
         escape = F,
         align = c("r"),
         caption = "Parking surpluses (on-street and off-street) for all English councils (£ millions)",
         linesep = "",
         col.names = c("\\multirow{1}{*}[0pt]{Local Authority}", yearz, 
                       paste0("Change ", current.year, "-", current.year-1999, 
                              " on ", current.year-1, "-", current.year-1999-1))) %>% 
  kable_styling(full_width = TRUE, 
                latex_options =c("HOLD_position", "striped", "repeat_header"), 
                font_size = 9) %>% 
  column_spec(1, width = "4.2cm") %>% 
  row_spec(c(353,355), hline_after = TRUE) 
```

\newpage
## References {-}
