---
title: "RAC Parking Report Automation---Queries for RAC, round 2"
author: mz
date: 19.3.2019
output:
  pdf_document: default
---

Previous queries: all good, thanks especially for the Nottingham data! Here's a new round, maily Scotland based. 

#### Scotland- existing reports 

The relevant data seems to be on this [Scottish local government finance statistics site](https://www2.gov.scot/Topics/Statistics/Browse/Local-Government-Finance/PubScottishLGFStats). See the table in the appendix, but I am confused by the following, although this is completely irrelevant right now:

9. Although there is a report for 2011/12, the tables on the website do not have individual LA level data, so I have no idea where the data came from?


#### Scotland-Available data

The available data is summarised in the appendix.

Because there were some obvious differences in the totals, I manually cross-checked the 2015/16 income and expenditure data vs the data published in the Leibling report. There are a few annoying discrepancies, like so:

```{r, echo = FALSE, messages = FALSE}
suppressMessages(library(dplyr))
library(knitr)
library(kableExtra)
account <- c("expednitures", "expednitures",
             "income", "income",
             "surplus")
la <- c("Aberdeen city", "South Ayrshire", 
        "Aberdeen city", "South Ayrshire", 
        "Aberdeen city")
leibling <- c(4320,470,9200,80,4890)
gov.scot <- c(0,-327,0,0,0)
df <- data.frame(account, la, leibling, gov.scot)
kable(df, "latex", booktabs = T, escape = F) %>%
  collapse_rows(columns = 1)
```

In fact Aberdeen City has zeros in all the Excel tables, but clearly there must be another source for this data. Apart from that there are just a few small rounding errors, so in the end after finding these big errors, everything else adds up OK. I don't want to go manually checking the rest, but I'm not sure what to do with this..

The Penalty Charge Notice data also has expenditures and incomes.But these apparently do not include off street parking, and possibly have other reasons to be different than the first set of tables. For 2016/17 the Aberdeen income data in Leibling is the same as the PCN one, but not the expenditure data. For previous years even that doesn't match. So still not sure where the source of Aberdeen City data is. 

10. Where can I get the Aberdeen City data?

11. Any idea what is wrong with the South Ayrshire data for 15/16? 

12. Apart from Aberdeen City, am I OK to use all other 31 authorities as they are reported in the Excel files?

On several occasions Leibling uses "Total transport expenditure" data, either directly or as a denominator. I cannot figure out what that is. In the Excel files it is not the total "Roads and Transport", nor is it the sum of "Roads and transport" minus parking, which he does in the England report. 

13. Any idea what Leibling is using as the total transport expenditure value? Or rather what do you want me to use?

14. Also, do you want to continue the comparison between the incomes and expenditures reported in both sets of files like Leibling did in the 2016/17 report (section 6)?






## England mapping

OK, so if I've got this right, the data is for 353 local authorities, of different types. This includes both 27 county councils and 201 district councils, however these both *cover the same area*. And both collect parking income. 

So how do you map that? When it comes to income, expenditure or surplus, i.e. ££, I guess you could e.g. take the county numbers and divide them equally between the districts? Equally though? Might there be a reason for a different type of division?

Alternatively I could add the sub-districts with the county total, and just plot those 27 + and then there are the other 8 counties that are not double counted, and merge them as well, so the England map has 35 LAs. But that's not consistent with the rest of the report..

15. Do you have a preference, alternative idea, suggestion? Otherwise there will be no England maps..

#### General aesthetic question

Some of the tables get pretty dense. At least in some cases (e.g. %) it might make sense to use conditional formatting to colour the cells. 

16. Would you be interested in that, I can give it a go?


## Appendix 1

### Scotland data - incomes and expenditures

```{r appendix, child = 'app3.Rmd'}
```

