
## England--Outturn
```{r, echo = FALSE}
library(kableExtra)
england.outturn <- readRDS(here::here("data/02-interim/england.outturn.17.18.rds"))
kable(england.outturn, "latex", booktabs = T, escape = F) %>%
  collapse_rows(columns = 1)
```

  
* *year*--period covered in report
* *link*--link to data file
* *rows*--number of rows (not all are local authorities though!)
* *las*--number of local authorities (i.e. not ones with class "O"
* *first*--first row of data on sheets 3,4 and 5
* *e.sh*--which sheet are expenditures on
* *e.on*--which column on street expenditures
* *e.off*--which column are off street expenditures
* *e.cc*--which column are congestion charge expenditures on (row = Greater London Authority)
* *i.sh*--which sheet are incomes on
* *i.on*--which column on street incomes
* *i.off*--which column are off street incomes
* *i.cc*--which column are congestion charge incomes on (row = Greater London Authority)
* *pen.sh*--which sheet are penalty charge incomes for LAs on
* *pen.on*--which column are penalty charge incomes for LAs
* *pen.1*--which cell is the totaL Penalty charge notice income for on street parking data on in sheet 2
* *tot.1*--which cell is the Net current expenditures for Total transport and Highways services on in sheet 2

## England--Budget
```{r, echo = FALSE}
england.budget <- readRDS(here::here("data/02-interim/england.budget.18.19.rds"))
kable(england.budget, "latex", booktabs = T, escape = F) %>%
  collapse_rows(columns = 1)
```

  
* *year*--period covered in report
* *link*--link to data file
* *rows*--number of local authorities
* *budg.tot*--wich cell on sheet 2 is the total budgeted surplus in
* *budg.la*--which column on sheet 3 are the LA parking budgeted surpluses on
