
```{r, echo = FALSE}
library(kableExtra)
scotland.i.e. <- readRDS(here::here("data/01-raw/orig.sco.meta.i.e.rds"))
kable(scotland.i.e., "latex", booktabs = T, escape = F) %>%
  collapse_rows(columns = 1)
```

* *year*--period covered in report
* *report*--does the report exist for this year?
* *file.name*--name on link to file on the website
* *link*--link to data file
* *start.sh*--number of first sheet with LA data
* *end.sh*--number of last sheet with LA data
* *exp.cell*--cell with expenditure data
* *inc.cell*--cell with income data
* *t.exp.cell*--total expenditure for all transport and roads cell
* *t.inc.cell*--total income for all transport and roads cell

#### DPE and PNC data

```{r, echo = FALSE}
scotland.pnc <- readRDS(here::here("data/01-raw/orig.sco.meta.pdf.rds"))

kable(scotland.pnc, "latex", booktabs = T, escape = F) %>%
  collapse_rows(columns = 1)
```

* *year*--period covered in report
* *report*--does the report exist for this year?
* *file.name*--name on link to file on the website
* *link*--link to data file
* *dpe.tab*--which table (including the table of contents) contains the table of LAs that have/haven't DPE
* *pcn.tab*--which table (including the table of contents) contains data on no of PCNs issued
* *i.e.tab*--which table (including the table of contents) contains data on incomes and expenditures of DPE LAs